Supplementary Materials And Methods

### 2019-12-06 NAR_form_article.Rmd
addition
06 dec dit on article

abstract: |
  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation. Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions. In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values. Our model is able to handle miRNA interaction factors such as seed type, binding energy, if provided. Our approach is able to reveal that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be calculated. Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies. We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs. We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods. A gene over-expressed in tumor tissue, namely *SERPINE2*, is used as perturbing element. We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context. Therefore, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account. The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs. The model is available as R package ceRNAnetsim https://github.com/selcenari/ceRNAnetsim.

@@

# INTRODUCTION

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms. Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets. Repressive activities of miRNAs vary depending on many factors that are significant to miRNA:target interactions. These factors include miRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007]. Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012]. It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012]. Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to specific sequence located in 3'UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005]. In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015]. These studies also showed that  binding location either indicate functionality of miRNA:target interaction or affect abundance of targets. It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007]. On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides [@chi2012alternative] and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013]. On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].

@@

# MATERIALS AND METHODS

@@

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=21cm,height=12cm]{fig1_new.eps}
\end{center}
\caption{Schematic presentation of mechanism of network based model. 
    (\textbf{A}) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
    (\textbf{B}) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
    (\textbf{C}) Distribution of miRNA1 (M1) changes. 
    (\textbf{D)}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2). Expression values are rounded to integers for simplicity. Gray values on edges indicate initial distribution, red values indicate increase and blue values indicate decrease in distribution. Shades of circles indicate different levels of increase.}
\label{fig:fig1}
\end{figure*}

@@

## Breast cancer patient dataset
We have applied our model in a real dataset for which experimental measurements of various factors were available. Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga. High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods). We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015] (Section 6 in Supplementary Materials and Methods and Table S4 in Supplementary Tables). SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network. Relevant steps and codes are available at Supplementary Materials and Methods.

# RESULTS

@@

\begin{table*}[ht]
\centering
\caption{Expression values of elements and interaction factors of miRNA:target interactions} 
\begin{tabular}{rllrrrrr}
  \hline
 & competing & miRNA & Competing expression & miRNA expression & seed type & region & energy \\ 
  \hline
   & Gene1 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.30 & -20.00 \\ 
   & Gene2 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.01 & -15.00 \\ 
   & Gene3 & Mir1 & 5000.00 & 1000.00 & 0.32 & 0.40 & -14.00 \\ 
   & Gene4 & Mir1 & 10000.00 & 1000.00 & 0.23 & 0.50 & -10.00 \\ 
   & Gene4 & Mir2 & 10000.00 & 2000.00 & 0.35 & 0.90 & -12.00 \\ 
   & Gene5 & Mir2 & 5000.00 & 2000.00 & 0.05 & 0.40 & -11.00 \\ 
   & Gene6 & Mir2 & 10000.00 & 2000.00 & 0.01 & 0.80 & -25.00 \\ 
   \hline
\end{tabular}
\label{tab:one}
\end{table*}

@@

# DISCUSSION

@@

# CONCLUSION

@@

\newpage

>>>>

abstract: |
  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation. Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions. In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values. Our model is able to handle miRNA interaction factors such as seed type, binding energy, if provided. Our approach is able to reveal that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be calculated. Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies. We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs. We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods. A gene over-expressed in tumor tissue, namely *SERPINE2*, is used as perturbing element. We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context. Therefore, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account. The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs. The model is available as R package ceRNAnetsim (https://github.com/selcenari/ceRNAnetsim).

@@

# \textbf{INTRODUCTION}

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms [@brennecke2005principles]. Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets. Repressive activities of miRNAs vary depending on many factors that are significant to miRNA:target interactions. These factors include miRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007]. Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012]. It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012]. Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to specific sequence located in 3'UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005]. In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015]. These studies also showed that  binding location either indicate functionality of miRNA:target interaction or affect abundance of targets. It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007]. On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides [@chi2012alternative] and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013]. On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].

@@

# \textbf{MATERIALS AND METHODS}

@@

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=15cm,height=8cm]{fig1_new.eps}
\end{center}
\caption{Schematic presentation of mechanism of network based model. 
    (\textbf{A}) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
    (\textbf{B}) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
    (\textbf{C}) Distribution of miRNA1 (M1) changes. 
    (\textbf{D}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2). Expression values are rounded to integers for simplicity. Gray values on edges indicate initial distribution, red values indicate increase and blue values indicate decrease in distribution. Shades of circles indicate different levels of increase.}
\label{fig:fig1}
\end{figure*}

@@

## Breast cancer patient dataset
We have applied our model in a real dataset for which experimental measurements of various factors were available. Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network; https://www.cancer.gov/tcga. High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods). We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015] (Section 6 in Supplementary Materials and Methods and Table S4 in Supplementary Tables). SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network. Relevant steps and codes are available at Supplementary Materials and Methods.

# \textbf{RESULTS}

@@

\begin{table*}[ht]
\centering
\caption{Expression values of elements and interaction factors of miRNA:target interactions} 
\begin{tabular}{rllrrrrr}
  \hline
 & Competing & miRNA & Competing Expression & miRNA Expression & Seed Type & Region & Energy \\ 
  \hline
   & Gene1 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.30 & -20.00 \\ 
   & Gene2 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.01 & -15.00 \\ 
   & Gene3 & Mir1 & 5000.00 & 1000.00 & 0.32 & 0.40 & -14.00 \\ 
   & Gene4 & Mir1 & 10000.00 & 1000.00 & 0.23 & 0.50 & -10.00 \\ 
   & Gene4 & Mir2 & 10000.00 & 2000.00 & 0.35 & 0.90 & -12.00 \\ 
   & Gene5 & Mir2 & 5000.00 & 2000.00 & 0.05 & 0.40 & -11.00 \\ 
   & Gene6 & Mir2 & 10000.00 & 2000.00 & 0.01 & 0.80 & -25.00 \\ 
   \hline
\end{tabular}
\label{tab:one}
\end{table*}

@@

# Additions to be fixed

Additionally, miRNA:gene expression dataset which was integrated via miRTarBase [@chou2017mirtarbase] was analysed in mean of perturbation efficiency to determine crucial genes on the network. We have observed that two miRNAs and 488 genes are effective in terms of perturbation efficiency, when we ran our model with more than 15% efficiency on target expression. miR-10b was investigated via HMDD database [@huang2018hmdd] and mostly found as metastatic biomarker. On the other hand, it was observed that miR-148a is critical for promotion of apoptosis and supression of breast cancer cell growth. Additionaly, to obtain the number of genes in a narrow range, the model was ran with 17% efficency on target expression and 297 genes and one miRNA were observed as highly perturbation-effective. We analysed these genes via eDGAR, a database of Disease-Gene Associations with annotated Relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.  With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes. 33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset. 44 of regulated genes via these TFs were found in our highly effective genes. it has been observed that these genes cover 22.5% of 297 perturbation-effective genes.

When we analysed these genes via biomaRt OMIM query, it has been observed that our model provides enriched 1.45 fold results with hypergeometric p-values 3.19e-05.

# \textbf{DISCUSSION}

@@

# \textbf{CONCLUSION}

@@

### 2019-12-06 NAR_form_article.tex
addition
06 dec dit on article

\usepackage{float}

\floatplacement{figure}{H}

\usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref} \usepackage[T1]{fontenc}

@@

\begin{abstract}
Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs. We applied the model to real sample consisting of
breast cancer gene and miRNA expression dataset and experimental
miRNA:target interaction dataset all generated via high throughput
sequencing methods. A gene over-expressed in tumor tissue, namely
\emph{SERPINE2}, is used as perturbing element. We have observed that
change in expression level of single gene in miRNA:target network is
sufficient to perturb regulations in whole network, due to unforeseen
and unpredicted regulation which are only visible when considered in
network context. Therefore, this model helps unveiling the crosstalk
between elements in miRNA:target network where abundance of target and
sponge effect are taken into account. The model is scalable and can be
plugged in with emerging miRNA effectors such as circRNAs. The model is
available as R package ceRNAnetsim
\url{https://github.com/selcenari/ceRNAnetsim}.
\end{abstract}

\section{INTRODUCTION}

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to miRNA:target interactions. These factors include miRNA:target binding
energy, binding location in target sequence, base pairing types between
miRNA and target, abundance of miRNAs and targets
\citep{grimson_microrna_2007}. Binding energies of miRNA:target
complexes vary based on nucleotide context and determine folding
stability of complex \citep{cao_predicting_2012}. It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex \citep{helwak_mapping_2013} and does
not directly determine repressive activity of miRNA
\citep{cao_predicting_2012}. Early studies have argued that 2-8 nt
sequence, seed, located in miRNA 5'end bind to specific sequence located
in 3'UTR of its target
\citep{bartel_micrornas_2004, lewis_conserved_2005}. In recent studies,
it has been shown that miRNAs can interact with targets via sequences
located in regions such as 5'UTR or CDS
\citep{hausser_analysis_2013, helwak_mapping_2013, moore_mirnatarget_2015}.
These studies also showed that binding location either indicate
functionality of miRNA:target interaction or affect abundance of
targets. It has been shown that miRNAs exhibit repressive activity via,
6-8 nt long sequence that is perfectly complementary with targets, seed
region at the 5' end of miRNAs
\citep{bartel_micrornas:_2009, grimson_microrna_2007}. On the other
hand, some researchers have reported that seed sequence of miRNA can
have mismatches or bulged/wobble nucleotides \citep{chi2012alternative}
and may locate in region other than 5'end of miRNAs
\citep{hafner_transcriptome-wide_2010, helwak_mapping_2013}. On top of
all these factors, abundance of miRNAs and targets and miRNA:target
ratio in cells predominantly affect efficiency of miRNA:target
interaction
\citep{arvey_target_2010, bosson_endogenous_2014, denzler_assessing_2014}.

@@

\section{MATERIALS AND METHODS}

@@

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from TCGA Research Network \url{https://www.cancer.gov/tcga}.
High-throughput experimental datasets which are provided miRNA:gene
target pairs with interaction factors
\citep{helwak_mapping_2013, moore_mirnatarget_2015} (Detailed codes are
available at Section 6 in Supplementary Materials and Methods). We have
combined miRNA and gene expression datasets via miRNA:target gene
dataset retrieved from published datasets
\citep{helwak_mapping_2013, moore_mirnatarget_2015} (Section 6 in
Supplementary Materials and Methods and Table S4 in Supplementary
Tables). SERPINE2, one of the most effective nodes, has been used as
trigger in ceRNA network simulation in breast cancer patient network.
Relevant steps and codes are available at Supplementary Materials and
Methods.

\section{RESULTS}

@@

\section{DISCUSSION}

@@

\section{CONCLUSION}

>>>>

\usepackage{float} \floatplacement{figure}{H} \usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref} \usepackage[T1]{fontenc}

@@

\begin{abstract}
Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs. We applied the model to real sample consisting of
breast cancer gene and miRNA expression dataset and experimental
miRNA:target interaction dataset all generated via high throughput
sequencing methods. A gene over-expressed in tumor tissue, namely
\emph{SERPINE2}, is used as perturbing element. We have observed that
change in expression level of single gene in miRNA:target network is
sufficient to perturb regulations in whole network, due to unforeseen
and unpredicted regulation which are only visible when considered in
network context. Therefore, this model helps unveiling the crosstalk
between elements in miRNA:target network where abundance of target and
sponge effect are taken into account. The model is scalable and can be
plugged in with emerging miRNA effectors such as circRNAs. The model is
available as R package ceRNAnetsim
(\url{https://github.com/selcenari/ceRNAnetsim}).
\end{abstract}

\section{\texorpdfstring{\textbf{INTRODUCTION}}{}}

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms \citep{brennecke2005principles}. Although the mechanisms by
which miRNA represses are not fully understood, miRNAs predominantly
repress their targets. Repressive activities of miRNAs vary depending on
many factors that are significant to miRNA:target interactions. These
factors include miRNA:target binding energy, binding location in target
sequence, base pairing types between miRNA and target, abundance of
miRNAs and targets \citep{grimson_microrna_2007}. Binding energies of
miRNA:target complexes vary based on nucleotide context and determine
folding stability of complex \citep{cao_predicting_2012}. It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex \citep{helwak_mapping_2013} and does
not directly determine repressive activity of miRNA
\citep{cao_predicting_2012}. Early studies have argued that 2-8 nt
sequence, seed, located in miRNA 5'end bind to specific sequence located
in 3'UTR of its target
\citep{bartel_micrornas_2004, lewis_conserved_2005}. In recent studies,
it has been shown that miRNAs can interact with targets via sequences
located in regions such as 5'UTR or CDS
\citep{hausser_analysis_2013, helwak_mapping_2013, moore_mirnatarget_2015}.
These studies also showed that binding location either indicate
functionality of miRNA:target interaction or affect abundance of
targets. It has been shown that miRNAs exhibit repressive activity via,
6-8 nt long sequence that is perfectly complementary with targets, seed
region at the 5' end of miRNAs
\citep{bartel_micrornas:_2009, grimson_microrna_2007}. On the other
hand, some researchers have reported that seed sequence of miRNA can
have mismatches or bulged/wobble nucleotides \citep{chi2012alternative}
and may locate in region other than 5'end of miRNAs
\citep{hafner_transcriptome-wide_2010, helwak_mapping_2013}. On top of
all these factors, abundance of miRNAs and targets and miRNA:target
ratio in cells predominantly affect efficiency of miRNA:target
interaction
\citep{arvey_target_2010, bosson_endogenous_2014, denzler_assessing_2014}.

@@

\section{\texorpdfstring{\textbf{MATERIALS AND METHODS}}{}}

@@

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from TCGA Research Network; \url{https://www.cancer.gov/tcga}.
High-throughput experimental datasets which are provided miRNA:gene
target pairs with interaction factors
\citep{helwak_mapping_2013, moore_mirnatarget_2015} (Detailed codes are
available at Section 6 in Supplementary Materials and Methods). We have
combined miRNA and gene expression datasets via miRNA:target gene
dataset retrieved from published datasets
\citep{helwak_mapping_2013, moore_mirnatarget_2015} (Section 6 in
Supplementary Materials and Methods and Table S4 in Supplementary
Tables). SERPINE2, one of the most effective nodes, has been used as
trigger in ceRNA network simulation in breast cancer patient network.
Relevant steps and codes are available at Supplementary Materials and
Methods.

\section{\texorpdfstring{\textbf{RESULTS}}{}}

@@

\section{Additions to be fixed}

Additionally, miRNA:gene expression dataset which was integrated via
miRTarBase \citep{chou2017mirtarbase} was analysed in mean of
perturbation efficiency to determine crucial genes on the network. We
have observed that two miRNAs and 488 genes are effective in terms of
perturbation efficiency, when we ran our model with more than 15\%
efficiency on target expression. miR-10b was investigated via HMDD
database \citep{huang2018hmdd} and mostly found as metastatic biomarker.
On the other hand, it was observed that miR-148a is critical for
promotion of apoptosis and supression of breast cancer cell growth.
Additionaly, to obtain the number of genes in a narrow range, the model
was ran with 17\% efficency on target expression and 297 genes and one
miRNA were observed as highly perturbation-effective. We analysed these
genes via eDGAR, a database of Disease-Gene Associations with annotated
Relationships among genes \citep{babbi2017edgar} that supplies different
observations about crucial genes such as direct physical interactions,
other shared genes with these interactions, protein-protein
interactions, GO terms etc. With our model, we obtained 3 (TP53, BRCA1,
AKT1) of 27 critical breast cancer genes and these genes show direct
physical interaction with 21 genes. 33 shared genes which are observed
on these direct interactions were also found in our effective
perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4,
GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical
TFs which are provided by eDGAR dataset. 44 of regulated genes via these
TFs were found in our highly effective genes. it has been observed that
these genes cover 22.5\% of 297 perturbation-effective genes.

When we analysed these genes via biomaRt OMIM query, it has been
observed that our model provides enriched 1.45 fold results with
hypergeometric p-values 3.19e-05.

\section{\texorpdfstring{\textbf{DISCUSSION}}{}}

@@

\section{\texorpdfstring{\textbf{CONCLUSION}}{}}

### 2019-12-06 Supplementary_mat_met.Rmd
06 dec dit on article

```{r, message=FALSE, warning=FALSE}
#install.packages("devtools")
#devtools::install_github("selcenari/ceRNAnetsim")
library(ceRNAnetsim)
```

@@

perturbationofnetwork%>%
  filter(type== "Competing")%>%
  arrange(desc(perturbation_efficiency))
```
This dataset, `perturbationofnetwork`, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.

>>>>

```{r, message=FALSE, warning=FALSE}
#install.packages("devtools")
#devtools::install_github("selcenari/ceRNAnetsim")
library(ceRNAnetsim)
library(stringr)
```

@@

perturbationofnetwork%>%
  filter(type== "Competing")%>%
  arrange(desc(perturbation_efficiency))
```
This dataset, `perturbationofnetwork`, includes totally 420 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
